Automated Topic Naming

### 2012-01-19 ese/ese.tex
some rewording in IRR section, migod's comment there

\subsection{Do NFR Frequencies Change over Time?}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

\subsection{Do Projects Differ in their Relative  Topic Interest?}
\emph{Yes}. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and	
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
PostgreSQL had more NFR topics in general, but \emph{portability},
\emph{functionality} and \emph{usability} were more prevalent in
PostgreSQL. One notable difference between PostgreSQL and the other
two projects is that many of the commits for PostgreSQL were UI
changes and documentation changes. PostgreSQL also seemed to focus
more on \emph{reliability} later in the case study than earlier like MySQL.
In all cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
how often \emph{portability} arises as a concern.

@@

\subsection{Do different developers work on different NFRs?}

@@

Our working theory is that less frequent committers are more
specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either ``wear more hats'' or have more
time to ``wear more hats''. An interesting extension would be to compare this data with 
the PostgreSQL source code files, to see which developer touched which file.

@@

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developer's likelihood to use or not use terms found within
our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
terms of number of commits, will have more samples, and thus be more likely
take on a more general role. For example, we found in a previous
project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
and the actual author of the commit. 
%were often correlated. Thus based
%on that observation developer style might be a confounding concern.

@@

\subsection{Annotation Observations}
We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
% management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes concerned updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.

@@

\subsection{Inter-rater reliability}
\label{sec:inter-rater}
To determine inter-rater reliability two of the authors---Ernst and Hindle---each annotated the PostgreSQL topics,
and then evaluated each other's annotations.

@@

Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
correlation of our per topic annotations for each NFR. We had to
evaluate inter-rater reliability this way because a single topic could be
tagged with more than one NFR.

We feel these results are poor. The aggregate view of a Kappa of
$0.1$ indicates there is some weak agreement. We found that
there was a lot of agreement in terms of lack of a tag, but often more
disagreement when a tag should be applied. After some discussion we concluded that usability was a source of
disagreement. Do user manuals improve usability? Is a command-line
option a usability issue? These kinds of questions illustrate some of
the agreement, disagreement and ambiguity about these labels. 
Thus we would recommend that future annotators make a concerted effort
to train and discuss when one applies an annotation.

Empirically how poor were our results? We decided to compare our
results against simulated or random results. In the first simulation
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling versus our labelling, and repeated
this 100,000 times, we found that
for 4 out of 7 NFR labels, our labels IRR were greater than the random
IRR measures 96\% of the time. For portability and efficiency, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
these sampled ratings versus our IRR using the Kappa statistic.

We also achieved similar results if the ratings were pulled from a
uniform distributions. What this does indicate is that our labels and
ratings were definitely better than random except for functionality,
reliability and usability.

@@

We feel that this resulting IRR score, although lackluster, indicates
the underlying difficulty of annotating such a data-set.
Consistent training of the annotators (the people) might avoid low IRR scores, or
using more annotators.
Other improvements could be gained by approaching some of the original developers
and using their expertise to annotate the data according to their intent.
Thus improvements
could be gained by producing more robust training samples that use
multiple raters, or experts, who are similarly trained and willing to discuss and
negotiate disagreements.

>>>>

\subsection{RQ1: Do NFR Frequencies Change over Time?}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

\subsection{RQ2: Do Projects Differ in Their Relative  Topic Interest?}
We found significant overall variation in the projects. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and	
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
PostgreSQL had more NFR topics in general, but \emph{portability},
\emph{functionality} and \emph{usability} were more prevalent in
PostgreSQL. One notable difference between PostgreSQL and the other
two projects is that many of the commits for PostgreSQL were UI
changes and documentation changes. PostgreSQL also seemed to focus
more on \emph{reliability} later in the case study than earlier like MySQL.
In all cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
how often \emph{portability} arises as a concern.

@@

\subsection{RQ3: Do Different Developers Work on Different NFRs?}

@@

Our working theory is that less frequent committers are more
specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either have wider responsibility or have more
time to be widely involved. An interesting extension would be to compare this data with 
the PostgreSQL source code files, to see which developer touched which file.

@@

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developer's likelihood to use terms found within
our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
terms of number of commits, will have more samples, and thus be more likely
take on a more general role. For example, we found in a previous
project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
and the author of the commit. 
%were often correlated. Thus based
%on that observation developer style might be a confounding concern.

@@

\subsection{Annotation Observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
Topics related to \emph{Configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
% management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes concerned updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.

@@

\subsection{Inter-rater reliability}
\label{sec:inter-rater}
To determine inter-rater reliability two of the authors---Ernst and Hindle---each annotated the PostgreSQL topics,
and then evaluated each other's annotations. 
Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
correlation of our per-topic annotations for each NFR. We 
evaluated inter-rater reliability using each NFR, because a single topic could be
tagged with more than one NFR.

@@

We feel these results are poor. The aggregate view of a Kappa of
$0.1$ indicates there is some weak agreement. We found that
there was a lot of agreement in terms of lack of a tag, but often more
disagreement when a tag should be applied. After some discussion we concluded that 
usability was a primary source of
disagreement. Do user manuals improve usability? Is a command-line
option a usability issue? These kinds of questions illustrate some of
the agreement, disagreement, and ambiguity about these labels. 
We therefore recommend that future annotators train and discuss how and when an annotation is appropriate.

To evaluate our results empirically, we compared them the annotations that would result if the seven labels
 were applied randomly. In the first simulation
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling, compared it with our labelling, and repeated
this 100,000 times. We found that
for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from a
uniform distributions. These experiments indicate that our labels and
ratings were an improvement on random annotations in all cases, with the exceptions of \emph{functionality},
\emph{reliability} and \emph{usability}.

@@

We feel that this resulting IRR score indicates
the underlying difficulty of annotating such a data-set. To improve IRR,
consistent training of the annotators or the use of more annotators might help.
Other improvements could be gained by approaching some of the original developers
and using their expertise to annotate the data according to their original, recollected intent.
Improvements
could be therefore be gained by producing more robust training samples that use
multiple raters, or experts, who are similarly trained and willing to discuss and
negotiate disagreements.
